---
title: 'Week 3: Assignment'
author: "Eric Lemphul"
date: "9/12/2021"
output: html_document
---

# Overview

Assignment 3 will cover topics related to data manipulation and regular expressions.

# Question 1:
## Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
# Get data from github
college.majors.data <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")

# subset data to only include DATA and STATISTICS
college.majors.subset <- subset(college.majors.data, grepl("DATA", Major) | grepl("STATISTICS", Major)) # grepl() returns TRUE if condition is met

print(college.majors.subset)
```


# Question 2:
## Write code that transforms the data below:

### [1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

### [5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

###  "elderberry"   "lime"         "lychee"       "mulberry"    

### [13] "olive"        "salal berry"

## Into a format like this:

### c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")


```{r}
# Data from above
fruit.data <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'

# split data by each element
fruit.data.scan<-scan(text=fruit.data, what='character', quiet=TRUE)

# remove the elements that are not in the new format (example: [1])
fruit.data.scan<-fruit.data.scan[-c(1,6,11,16)]

# prints the data in the desired format
fruit.data.reformat<-dput(fruit.data.scan)
```

# Question 3:
## Describe, in words, what these expressions will match:




### (.)\1\1

(.)\1\1 will match any string that repeats back to back. Example: aaa


### "(.)(.)\\\2\\\1"

"(.)(.)\\\2\\\1" will match to a symmetrical group of 4 characters. Example: abba

### (..)\1


(..)\1 will match any string that has 2 characters that repeat once. Example: abab

### "(.).\\\1.\\\1"

"(.).\\\1.\\\1" will match any string that repeats the same character 3 times, but in between each repeated character there is a random character. Example: abaca

### "(.)(.)(.).*\\\3\\\2\\\1"

"(.)(.)(.).*\\\3\\\2\\\1" will match with 6 characters that are symmetric with any set of values in between the symmetry. Examples: "abccba"   "abc33345cba"


# Question 4:
## Construct regular expressions to match words that:

```{r}
# sample words
words = c("eye", "noon", "harsh", "church", "shush", "eleven", "banana", "eevee")
```

### Start and end with the same character.

The regular expression is: ^(.)((.*\\\1\$)|\\\1?\$)
```{r}
# Regular Expression: ^(.)((.*\\1$)|\\1?$)
grep("^(.)((.*\\1$)|\\1?$)",words,value=T)
```

### Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

The regular expression is: ([A-Za-z][A-Za-z]).*\\\1
```{r}
# Regular Expression: ([A-Za-z][A-Za-z]).*\\1
grep("([A-Za-z][A-Za-z]).*\\1",words,value=T)
```

### Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

The regular expression is: ([A-Za-z]).\*\\\1.\*\\\1
```{r}
# Regular Expression: ([A-Za-z]).*\\1.*\\1
grep("([A-Za-z]).*\\1.*\\1",words,value=T)
```
