---
title: "Assignment - SQL and R"
author: "Eric Lemphul"
date: "9/5/2021"
output: html_document
---

# Overview

This assignment required me to collect data about movie ratings.  The six movies that I chose to collect data on were Greyhound, Disney: Jungle Cruise, Don't Breathe 2, Free Guy, Orphan, and Black Widow. I used MySQL as the database for this assignment.


# MySQL 

The first task I completed was creating the MOVIE table in MySQL. After creating the table, I used the MySQL Workbench Table Data Export Wizard to easily export the table as a .csv file.  To export using the Wizard, first right click on the table in the schema after creating the table. Next, select all variables in the table. Next, chose the file path (I chose my R project file), add the name f the file (I named it movie.csv), and make sure that it will export as a .csv file to local machine. I kept the Line Separator option as LF, the enclose string as ",I changed the Field Separator to ,(comma), and I kept the null and NULL as SQL keywords as YES. Continue clicking next to execute the exportation of the MySQL table as a .csv file.

The MySQL code that I used to create the table is in a .sql file in my github.


# R code


Now that the data has been sucessfully exported, it is time to read the .csv file.
```{r}
movie.data<-read.csv("movie.csv")
movie.data
```

## Data Preprocessing


Let's handle the missing values. (I entered the values as NULL in the .sql script, but they were entered in as 0, 0.0 or "" in the R data frame)


First let's find which movie the missing data in row 22 is for.

```{r}
movie.data[22,]

sum(movie.data$Movie == "Greyhound")
sum(movie.data$Movie == "Disney: Jungle Cruise")
sum(movie.data$Movie == "Don't Breathe 2")
sum(movie.data$Movie == "Free Guy")
sum(movie.data$Movie == "Orphan") # Orphan only has 4 entries where every other entry has 5
sum(movie.data$Movie == "Black Widow")

movie.data$Movie[movie.data$Movie == ""]<-"Orphan"
movie.data$Director[movie.data$Director == ""]<-"Jaume Collet-Serra"
movie.data$MovieLengthMinutes[movie.data$MovieLengthMinutes == 0]<-123
movie.data$Genre[movie.data$Genre == ""]<-"Horror"

movie.data[22,]
```


Next let's deal with the missing values in the MovieRating column. I am setting all missing values equal to the mean of the movie rating.

```{r}
movie.data[,1:2]

movie.data$MovieRating[movie.data$Movie == "Greyhound" & movie.data$MovieRating == 0.0]<-mean(movie.data$MovieRating[movie.data$Movie == "Greyhound"])
movie.data$MovieRating[movie.data$Movie == "Disney: Jungle Cruise" & movie.data$MovieRating == 0.0]<-mean(movie.data$MovieRating[movie.data$Movie == "Disney: Jungle Cruise"])
movie.data$MovieRating[movie.data$Movie == "Don't Breathe 2" & movie.data$MovieRating == 0.0]<-mean(movie.data$MovieRating[movie.data$Movie == "Don't Breathe 2"])
movie.data$MovieRating[movie.data$Movie == "Free Guy" & movie.data$MovieRating == 0.0]<-mean(movie.data$MovieRating[movie.data$Movie == "Free Guy"])
movie.data$MovieRating[movie.data$Movie == "Orphan" & movie.data$MovieRating == 0.0]<-mean(movie.data$MovieRating[movie.data$Movie == "Orphan"])
movie.data$MovieRating[movie.data$Movie == "Black Widow" & movie.data$MovieRating == 0.0]<-mean(movie.data$MovieRating[movie.data$Movie == "Black Widow"])

movie.data[,1:2]
```


Let's factor the categorical data.

```{r}
movie.data$Movie<-factor(movie.data$Movie)
movie.data$Director<-factor(movie.data$Director)
movie.data$Genre<-factor(movie.data$Genre)
movie.data$LastName<-factor(movie.data$LastName)
movie.data$FirstName<-factor(movie.data$FirstName)
```


Summary of the pre-processed data frame


```{r}
summary(movie.data)
```


# Conclusions


The MOVIE data table that I created in MySQL Workbench required some data pre-processing after exporting it to a .csv file using the MySQL Workbench Table Data Export Wizard. NULL values from the movie.csv I created were dealt with accordingly. NULL values were made into 0, 0.0, or "" depending on the data type of the column. Numerical NULL values were imputed with the mean of their corresponding movie. The missing information in row 22 was dealt with using the sum function to determine which movie occurred the least in the data frame and was replaced with the proper information. 





